### 2020-10-19 Chapter3/3_6_conclusions.tex

\section{Conclusions}
\label{sec:3_6_conclusion}

In this chapter, I presented some analyses made possible through \tool, a platform I designed and described in chapter \ref{chap:2_dataset} to collect and store data, and able to extract higher level information from FFCS provider.

By analysing the data, I highlighted different aspects related to the system utilization, how users move in the city in different periods of the day, and what are the users' driving habits.

This study points out the big amount of information that it is possible to extract from well designed data collection pipelines. More in details, by looking the system utilization, I demonstrated that FFCS cars are frequently used for short trips which last less then 30 minutes and 5 km. Moreover, despite Enjoy has a smaller fleet, its system utilization is frequently higher than car2go one due to the more appreciated car model it offers. Exploiting the spatial analysis, I highlighted how users tend to move during different time periods. Finally, the users' driving habits showed us that current charging policy may encourage users to drive fast.

prova di commit