### 2020-10-19 Chapter7/letter.tex

\documentclass{letter}

%\setlength{\parskip}{0.11cm plus1mm minus1mm}
%\setlength{\paperheight}{11in}
%\setlength{\paperwidth}{8.5in}
%\usepackage[top=0.70in, bottom=0.70in, left=0.75in, right=0.75in]{geometry}

\begin{document}

Dear Editor and Reviewers,

This manuscript is a revised version of paper N. PMC\_2018\_465.
First, let us thank the Editor and the Reviewers for the useful comments.
We did our best to address the issues pointed out during the reviewing process. 
Please find attached our detailed reply to all the questions and comments raised during the reviewing process. We also attach as a supporting file a version of the paper ('diff.pdf') with all  the modifications highlighted in blue.

Luca Vassio (on behalf of the authors)

\end{document}