### 2020-11-17 thesis.tex

% !TEX encoding = UTF-8 Unicode
% !TEX TS-program = lualatex
% !BIB TS-program = biber

\documentclass[%
   corpo=12pt, % optional; default:= 10pt
   oneside, % recommended
   tipotesi=scudo,
   mybibliostyle, % only if biblio is typeset with a personal style
  numerazioneromana, % not necessary in a real thesis
   ]{toptesi}

\begin{filecontents*}{\jobname.xmpdata}
\Author{Michele Cocca}
\Title{Conversion of FFCS ICE fleet to EV: a data-driven approach}
\Subject{Doctoral dissertations in the SCUDO doctoral school}
\Keywords{PDF\sep
          PDF/A\sep
          ISO 19005\sep
          LaTeX\sep
          PhD Thesis\sep
          Engineering\sep
          SCUDO}
\Publisher{Politecnico di Torino}
\end{filecontents*}
%%%%%%% -----------------------------------------------------------

%%%% Use the following package if and only if you want to produce
%%%% an archivable document according to standard PDF/A-1b
%%%% No need to load package imakeidx, because it has already been
%%%% loaded by the specific module toptesi-scudo.
\usepackage[a-1b]{pdfx}
\usepackage[linesnumbered,ruled,norelsize]{algorithm2e}

\newtheorem{definition}{Definition}
%%%%%% Read the English documentation of TOPtesi in order to check
%%%%%% the special attention needed to produce ISO compliant
%%%%%% archivable files
%%%%

\ifPDFTeX
    \usepackage[utf8]{inputenc}% 
    \usepackage[T1]{fontenc}
\fi

\errorcontextlines=9% more information on the console in case of errors

%%%% Specify fonts here; chose one among these fonts by leaving just
%%%% one line without initial comment character.
%%%% With LuaLaTeX or XeLaTeX don't change fonts
\ifPDFTeX % using pdflatex
    \usepackage{lmodern} % Default
    %\usepackage{newtxtext,newtxmath}% Times eXtended for text and math
    %\usepackage{fourier}% Utopia, Helvetica and "monospace = ?"

\else % using lualatex (or xelatex)
% Here we use the Libertinus serif, sans serif, monospaced and math fonts;
% they are alla available with a complete up-to-date TeXLive installation.
% Without specifying any OpenType font, the Latin Modern OpenType ones are
% used by default; try commenting the setting for Libertinus Mono, run with
% LuaLaTeX and see what happens; you might prefer to keep the comment sign
% in this line.
    \usepackage{fontspec}
    \defaultfontfeatures{Ligatures=TeX}
    \setmainfont{Libertinus Serif}
    \setsansfont{Libertinus Sans}
    \setmonofont{Libertinus Mono}[Scale=MatchLowercase]
    \usepackage{unicode-math}% add special math stile option here
                             % for example [style=ISO]
% define one math font 
    \setmathfont{Libertinus Math}%
\fi

\usepackage{kantlipsum,mwe} % to produce dummy text and dummy figures
\usepackage{booktabs}

\makeindex[intoc]% collect material to index

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This conditional code is an example of using a different bibliography
% style from the one predefined by the class.
% The usage of the conditional code (also with a different contents) is
% discouraged, because the predefined numbered bibliography style is the
% one generally used in scientific publications.
% Even if discouraged, it is not forbidden.
\ifmybibstyle 
  \usepackage[autostyle]{csquotes} % necessary for biblatex
  \usepackage[backend=biber,
              style=philosophy-classic,
              scauthors=all,
              sorting=nyt,
              natbib]{biblatex} % LaTeX specific bibliography handler
\fi 
\addbibresource{\jobname.bib}% bibliographic data base(s)
% It is recommended to name a single or the primary .bib file with the same
% name as the thesis master file: the macro \jobname contains that name.
% It is possible lo use a comma separated list of bibliographical databases
% but extreme care must be paid to the fact that all .bib files have
% actually different names, and that the citations key are really distinct
% among alla databases, not simply within a single database.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The following is to be loaded as the end of the preamble if one wants
% to use hyperlinks and/or urls to be typed within the thesis,
% except that after loading hyperref very few commands may be
% issued. One is the \includeonly command with its list of files;
% other packages may be loaded after hyperref, only if their
% documentation says so; some of these critical packages, but they are
% not the only ones, involve Right to Left languages or other
% oriental languages.
%
% Distinguish the hyperref call from the hyperref setup so as
% to avoid option clashes with other packages that may invoke 
% hyperref with different options.

\unless\ifcsname ver@hyperref.sty\endcsname\usepackage{hyperref}\fi
\hypersetup{%
    pdfpagemode={UseOutlines},
    bookmarksopen,
    pdfstartview={FitH},
    colorlinks,
    linkcolor={blue},
    citecolor={blue},
    urlcolor={blue}
  }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%% The \includeonly argument should preferably be written with
%%%% one file name per line, so that by commenting or uncommenting
%%%% some lines a selective compilation may be executed.
\includeonly{%
Chapter1/1_introduction,%
Chapter2/2_datacollection,%
Chapter3/3_datacharact,%
Chapter4/4_cs_comparison,%
Chapter5/5_simulator,
Chapter6/6_4cities,
Chapter7/7_cs_optimization,
Chapter8/8_prediction,
%Chapter9/9_relocation,
Chapter10/10_scalability,
Chapter11/11_conclusion
%Appendix1/appendix1,
%Appendix2/appendix2,%
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\tool}{\textit{UMAP}\xspace}
\newcommand{\MGM}[1]{#1}
\newcommand{\removelatexerror}{\let\@latex@error\@gobble}
\newcommand{\reviewed}[1]{{\color{black}{#1}}}
\newcommand{\mc}[1]{{\color{purple}{[mike: #1]}}}
\usepackage[thinlines]{easytable}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\ifPDFTeX \usepackage{indentfirst}\fi
\raggedbottom

\begin{document}

% The contents of this ThesisTitlePage environment may be written
% in a configuration file named exactly the same as the thesis main
% file, but with extension .cfg. If similar commands with different
% data are written within this environment, such data prevail on
% those read from the configuration file.
\begin{ThesisTitlePage}
% Use the optional command to set a different School logo
% Its is possible to used this command several times; each time
% a new different  Institution logo  may be added. In general
% just the ScuDo logo is sufficient; for dissertations made in
% cooperation with the University of Turin, its logo may be added
% with a second instance of the \PhGschoolLog statement. With
% dissertations supported by the INRiM, this institution logo may
% be added. Such logos (in PDF format) may be obtained from the
% ScuDo web server.
\PhDschoolLogo{scudo}% Fake logo for this example
%\PhDschoolLogo{Logo-ScuDo-blu} % just the ScuDo logo
%\PhDschoolLogo{Logo-ScuDo-blu,Logo-INRIM} %for dissertations made in cooperation with INRIM
%\PhDschoolLogo{Logo-ScuDo-blu,Logo-INFN} %for dissertations made in cooperation with INFN
%\PhDschoolLogo{Logo-ScuDo-blu,Logo-UniTo} %for dissertations made in cooperation with the University of Turin
% Doctorate course name; mandatory
\ProgramName{Electronic and Telecommunication Engineering}
% Cycle ordinal number; optional.
% You can write 29.th, or 29\ap{th}, or 29\textsuperscript{th}, or ...
\CycleNumber{33.th}
% PhD candidate name; mandatory
\author{Michele Cocca}
% Dissertation title; mandatory
\title{Electric Revolution and\\Free Floating Car Sharing}
% Dissertation subtitle: optional.
% It might be useful only if the actual full title is too long.
\subtitle{A Data Driven Methodology for System Design}
% The supervisor(s) label; optional; default value "Supervisor:".
% You can change it to plural as in this example, where the colon has
% been eliminated.
%\NSupervisor{Supervisor}{Supervisors}
%
% The SupervisorNumber may contain a value such as 0, 1, or any
% number higher than 1. If 0 is specified, no label is typeset
% over the supervisor(s) list; if 1 is specified then the singular
% form is used: if any value higher than 1 is specified, the plural
% form is used.
\SupervisorNumber{1}
% List of supervisors with academic title, name(s), surname(s),
% and function; mandatory
\SupervisorList{%
    Prof.~Marco Mellia, Supervisor\\
    }
% Name of the examining committee: optional. 
% Default value "Doctoral Examination Committee"
%\Nexaminationcommittee{Doctoral examination committee}
% List of the  examining committee members: mandatory if the above label
% is not empty.
\ExaminerList{%
Prof.~A.B., Referee, University of \dots\\
Prof.~C.D., Referee, University of \dots\\
Prof.~E.F., University of \dots\\
Prof.~G.H., University of \dots\\
Prof.~I.J., University of \dots}
% Name of the institution where the examination takes place; optional.
% Default value: "Politecnico di Torino"
%\Nlocation{Politecnico di Torino}
% Examination date: mandatory, although the way to write it is free.
\ExaminationDate{\today}
% Disclaimer with signature; optional. Default text as
% in the following lines. 
\Disclaimer{%
\noindent I hereby declare that, the contents and organisation of this dissertation constitute my own original work and does not compromise in any way the rights of third parties, including those relating to the security of personal data.	
}
%\Signature{%
%\begin{flushright}
%\parbox{0.5\textwidth}{\centering
%\dotfill\\
%Mario Rossi\\
%Turin, February 29, 2123
%}
%\end{flushright}}
\end{ThesisTitlePage}
%%%%%%%%%%%%%%%% Everything else necessary in the thesis title
%%%%%%%%%%%%%%%% page and in the copyright page is supplied by
%%%%%%%%%%%%%%%% the default values.
%%%%%%%%%%%%%%%% If you enter an explicit disclaimer, you can
%%%%%%%%%%%%%%%% typeset other material before the disclaimer
%%%%%%%%%%%%%%%% formula; use the necessary spacing on order
%%%%%%%%%%%%%%%% to separate the formula from other text. 
%%%%%%%%%%%%%%%% 
%%%%%%%%%%%%%%%% For example you might want to write the formal
%%%%%%%%%%%%%%%% statement for a particular licence, provided the 
%%%%%%%%%%%%%%%% licence allows Open access; not necessarily all uses
%%%%%%%%%%%%%%%% of the thesis should be allowed, but the minimum
%%%%%%%%%%%%%%%% is the reading access.

%%%%%%%%%%%%%%%% The next two lines are metadata for a normal PDF file.
%%%%%%%%%%%%%%%% For ISO archivable PDF/A-1b metadata, they must
%%%%%%%%%%%%%%%% be entered only in the form used in the file
%%%%%%%%%%%%%%%% named in the filecontents* environment as shown
%%%%%%%%%%%%%%%% at the beginning of this file.

%\subject{How to typeset a doctoral thesis suitable for defence in almost any country and university.}
%\keywords{{pdfLaTeX} {LuaLaTeX} {XeLaTeX} {PhD doctoral programs} {PhD dissertation} {Politecnico di Torino}}

\summary%\sommario
%This is where you write your abstract \dots\ (Maximum 4000 characters, i.e. maximum two pages in normal sized font, typeset with the thesis layout).
%The abstract environment is also available, but  \texttt{\string\summary} is preferred because it generates an un-numbered chapter. The abstract environment is more suitable for articles and two column typesetting without a separate title page.
\mc{intro}

In my thesis, I presented a methodology to address, in different cases of studies, all the challenges related to the conversion of combustion engine cars to electric vehicles in FFCS. In particular, the main drivers focus on a profitable and technically sustainable system setup able to guarantee a flexible and appealing mobility service to an increasing customer audience.

In the first part of my thesis, I described the software I developed in order to scrape from the web real combustion engine FFCS, from two providers: car2go and Enjoy. The data collection car2go data collection lasted from \mc{december 2016 to january 2018, collecting more than 28 million users' bookings spread in 25 cities} while the Enjoy data collection phase started \mc{may 2018 and lasted until BOH}, collectin about \mc{BOH} bookings in 6 cities.

Then, I characterized both datasets in Turin, one of the cities in which both FFCS providers work. Initially, I defined a filtering criterion to remove outliers from the dataset and identify only real users' ride: entries reporting a trip where the user realistically moved. These analyses allowed me to identify geo-spatial user's patterns. After that, I compared the car2go customer's pattern with the one-way and two-way car-sharing system. The results showed how users prefer more flexible services like FFCS or one-way car sharing.

Once the data are consolidated, I developed (i) a methodology to assign a charging station to a city by looking at users' pattern; (ii) system policies to manage the fleet when the vehicle SoC may not guarantee a trip; (iii) an event-based trace-driven simulator able to replicate the recorded trips in an electrified scenario evaluating the feasibility of each configuration.

At this point I studied in Berlin, Milan, Turin, and Vancouver (cities having different traffic patterns) different electric FFCS setup. The best logic to place a charging station is to cover the zones recording more parkings: increasing the charging station presence in those zones, the probability to find a free plug increases too. Then, I showed how an electric FFCS is sustainable with few charging stations (8-10~\% of city zones) if the users accept to park the car in a charging station when their ride finishes nearby or when the SoC goes below a security threshold. Finally, I investigate different charging station setups. The best results come by balancing the spread of charging stations: concentrated all plugs in a unique hub or spread them all above the city leads to poor performances in terms of system sustainability and user's discomforts.

To still reduce the number of charging stations to have a sustainable electric FFCS, I compared several optimization algorithms. My results showed how a Genetic Algorithm is able to find a better solution able to shrink the minimum amount of resources to sustain the same mobility demand.

After that, I moved my attention to the users' rentals predictability. The main goal is to understand how different open-data sources could impact on the recorded FFCS users' rental. I used as a case of study the city of Vancouver, augmenting the rentals dataset with the weather and socio-economic open data. Initially, I compared several time-series forecasts to predict the users' demand in the short and medium-term. The Random Forrest regression produced better accuracy and results interpretability. Then I correlated the socio-economics features characterizing each Vancouver neighborhood to FFCS demand and again, the Random Forrest regression clearly showed a subset of socio-economic features more correlated to the FFCS users' rentals.

Finally, I questioned the system scalability figuring out several scenarios having increasing demand. I used a model able to synthesize users' demand by looking only at the geo-spatial users' rentals. By varying the electric FFCS setup and simulating the new scenario I pointed out how a linear increase in the demand intensity leads to a fleet sub-linear increase. Finally, I projected those considerations in euros proofing how electric FFCS has rooms for economic growth.

\begin{dedication}

\textit{I would like to dedicate this thesis to my loving parents}

{\normalsize 
The dedication very seldom is a proper thing to do; in some countries it is very common, while in other countries it is done for imitation of other people habits.

The sentence used above clearly is an example of something very common, but it is  useless. Of course we all love our beloved parents, but it is not necessary to ``engrave it in stone''.
\par}
\end{dedication}
%\end{dedica}

%%%%%%%%% If you want these two lists, uncomment the following line
%\tablespagetrue\figurespagetrue %

%%%%%%%%% Table of contents and optionally the tables and figures lists
\allcontents

\mainmatter %all the above is front matter; here begins the main matter

\include{Chapter1/1_introduction}
\include{Chapter2/2_datacollection}
\include{Chapter3/3_datacharact}
\include{Chapter4/4_cs_comparison}
\include{Chapter5/5_simulator}
\include{Chapter6/6_4cities}
\include{Chapter7/7_cs_optimization}
\include{Chapter8/8_prediction}
%\include{Chapter9/9_relocation}
\include{Chapter10/10_scalability}
\include{Chapter11/11_conclusion}
%% Numbered appendices remain in the main matter...
%\appendix
%\include{Appendix1/appendix1}
%\include{Appendix2/appendix2}

\backmatter% here begins the back matter
% ... otherwise a single appendix may stay here
%\include{References/biblio}

% Do not use this command if you did not prepare a nomenclature
% database by means of the suitable \nomenclature command and its
% arguments, as we did in chapter 2 of this example thesis.
\printnomencl

% In this example we use \nocite{*} in order to typeset the whole
% contents of the bibliographic database. Normally this is not
% necessary and it's better to let biber extract from the database
% only the cited works
%\nocite{*}

\printbibliography[heading=bibintoc]

% Do not use this command if you did not set the \makeindex switch
% in the preamble.

\printindex

\end{document}